### 2021-07-05 docs/manuscript/manuscript.tex
Rephrase text

As a measure of network complexity, we used the complexity index based on
Shannon entropy proposed by Bonchev \& Rovray~\cite{D.2005}. We calculated this
index using the following equation:
\begin{equation}
 I_{vd} = \sum_{i=1}^V a_i \cdot \log_{2} a_i,
\end{equation}
\noindent
being $a_i$ the degree of \emph{i}-th node and $V$ the number of nodes in the
network. The network descriptors were calculated using built-in functions from
Wolfram Mathematica software.\\

We performed a Pearson’s chi-square goodness of fit test for the node degree
distribution using Wolfram Mathematica.

\subsection*{Pattern formation under varying size of the morphogenetic field}
We wanted to evaluate if the GRNs evolved in our model with a morphogenetic
field size of 30 cells could produce the desired pattern in morphogenetic fields
with a different number of cells. For this reason, we calculated the fitness of
each GRN in morphogenetic fields with 10, 20, 40, and 50 cells. In these
simulations we redefined the morphogen gradient in such a way that the initial
and final concentration were constant in all the fields (\nameref{S3_Fig}A).
In order to know if topologies with the same topology had a similar fitness
in the different fields, we performed a Principal Component Analysis in
Scikit-learn~0.24.2~\cite{sklearn} using the fitness values of each GRN in
the four different fields.

@@

The classification of GRNs showed that our initial set of 2061 GRNs can be
grouped in a number of 714 distinct network topologies, each one of them
containing a different number of GRNs (Fig~\ref{fig:distopol}). This abundances
distribution was not produced by chance (Shannon entropy = 8.52963, 95\%
confidence interval = [10.8915, 10.9275]), and instead it displays some level
of structure, indicating the existence of network motifs.

@@

The neutral network (a metagraph in which each node is a network topology)
shows that most of the network topologies are grouped into a connected graph.
This graph is composed of 639 nodes that are accessible by one mutational step.
The 15 most abundant network topologies, with the exception of
number 10, are part of this graph (Fig~\ref{fig:neutral-network}). We also
found nodes inside the neutral network that had more connections between them
than with other nodes, generating clusters of nodes, interestingly in
almost every cluster is present one or more of the 15 most abundant network
topologies. In addition, we found that the first eight most abundant network
topologies were located in the largest cluster.

@@

Next to topological robustness, we investigated the robustness of the sampled
network topologies to parameter changes (see Methods). Fig~\ref{fig:ab-rob}A
illustrates the association between the abundance of a topology class versus its
robustness. We did not find a significant correlation between
these two variables ($r_S = −0.0389724$, $\text{p-value} = 0.298475$).
However, when we grouped those topologies containing
the same subgraph and calculated their mean robustness, we found that between
the most robust subgraphs is the ``Bistable'' motif reported by Cotterell \&
Sharpe~\cite{Cotterell2010}, the I3-FFL and the ``Overlapping domains'' motif
(which is a variation of the I3-FFL) also reported by Cotterell \& Sharpe
(see \nameref{S2_Fig} and \nameref{S1_Table}).

@@

Among the new network topologies we report here, perhaps one of the most
interesting ones is the topology number 9, which is able to express the striped
gene expression pattern with only two genes (Fig~\ref{fig:topol9}).
This topology is composed of a core that consists in the
Gierer-Meinhardt network~\cite{giererMeinh1972, meinhardt2000} known to
generate stripes and dots in bidimensional fields, and additional
interactions such as the self inhibition of gene A and the inhibition of the genes
by the morphogen. The Gierer-Meinhardt network consists in an activator of low
diffusivity (gene C) and an inhibitor with greater diffusivity (gene A). The additional
interactions of topology number 9 have the role of restricting the generation of
the expression band to a precise location.

>>>>

As a measure of network complexity, we used the complexity index based on
Shannon entropy proposed by Bonchev \& Rovray~\cite{D.2005}. We calculated this
index using the following equation:
\begin{equation}
 I_{vd} = \sum_{i=1}^V a_i \cdot \log_{2} a_i,
\end{equation}
\noindent
where $a_i$ denotes the degree of the \emph{i}-th node and $V$ the number of
nodes in the network. Both the network descriptors and Pearson’s chi-square
goodness of fit tests were performed for the node degree distribution using
the built-in functions from the Wolfram Mathematica software.\\

\subsection*{Pattern formation under varying size of the morphogenetic field}
We evaluated the ability of each of the sampled GRNs to generate a striped
pattern of gene expression under varying sizes of the morphogenetic field.
Specifically, we calculated the fitness of each GRN across morphogenetic fields
composed of 10, 20, 40, and 50 cells. In these simulations we redefined the
morphogen gradient in such a way that the initial and final concentration were
constant in all the fields (\nameref{S3_Fig}A). In order to know if topologies
with the same topology had a similar fitness in the different fields, we performed
a Principal Component Analysis in Scikit-learn~0.24.2~\cite{sklearn} using the
fitness values of each in the four different fields.

@@

The classification of GRNs showed that our initial set of 2061 GRNs can be
grouped into 714 distinct network topology classes, with each containing a
variable number of GRNs (Fig~\ref{fig:distopol}). This abundances
distribution was not produced by chance (Shannon entropy = 8.52963, 95\%
confidence interval = [10.8915, 10.9275]), and instead it displays some level
of structure, indicating the existence of network motifs.

@@

The neutral network (a metagraph in which each node is a network topology)
shows that most of the network topologies are grouped into a connected graph.
This graph is composed of 639 nodes that are accessible by one mutational step.
The 15 most abundant network topologies, with the exception of
number 10, are part of this graph (Fig~\ref{fig:neutral-network}). We also
found nodes inside the neutral network that had more connections between them
than with other nodes, which leads to further clustering of nodes. Interestingly in
almost every cluster one can find one or more of the 15 most abundant network
topologies. In addition, we found that the first eight most abundant network
topologies were located in the largest cluster.

@@

Next to topological robustness, we investigated the robustness of the sampled
network topologies to parameter changes (see Methods). Fig~\ref{fig:ab-rob}A
illustrates the association between the abundance of a topology class versus its
robustness. We did not find a significant correlation between
these two variables ($r_S = −0.0389724$, $\text{p-value} = 0.298475$).
However, when we grouped those topologies containing
the same subgraph and calculated their mean robustness, we found that between
the most robust subgraphs is the ``Bistable'' motif, the I3-FFL, as well as
the ``Overlapping domains'' motif (which is a variation of the I3-FFL), which
have been previously reported in Cotterell \& Sharpe~\cite{Cotterell2010}
(see \nameref{S2_Fig} and \nameref{S1_Table}).

@@

Among the new network topologies we report here, perhaps one of the most
interesting ones is the topology number 9, which is able to express the striped
gene expression pattern with only two genes (Fig~\ref{fig:topol9}).
This topology is composed of a core quite similar to the Gierer-Meinhardt
mechanism~\cite{giererMeinh1972, meinhardt2000} known to generate
stripes and dots in bidimensional fields, with the only differences being
the self inhibition of gene A and the inhibition of the genes by the morphogen.
Such additional interactions in the topology number 9 play a role in restricting
the striped expression pattern to a precise location along the morphogenetic field.
